Beyond clines: lineages and haplotype blocks in hybrid zones

### 2016-03-04 drafts/hybrid_zone.tex
.

\subsection*{The number of `unfit' ancestors}

\plr{This needs a bit more precise discussion.}

It is possible that there is additional information contained in patterns of within-ancestry haplotype diversity throughout the hybrid zone. Looking at the number of ancestral haplotypes present in the population would provide information about whether the migrant contribution is through relatively few, successful migrants, or through many migrants that each contribute relatively little.

We find that the average family size of a $B$ allele ($F_B(m,x)$) decrease with distance into the $A$ side of the hybrid zone, and is lowest far away from the zone center, where ancestry $B$ is at low frequency (Figure~\ref{Fig:family_size}). As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality. This pattern is expected and consistent with the prediction in our models that unfit lineages tend to be recent migrants, which will have smaller families. In conjunction with block length, measuring $F_B$ can provide an additional signal for characterizing lineages strongly influenced by selection.

%\alisa{add motivation-- additional information in haplotype diversity. Present alternate hypotheses. This ties in with our world view within our model}
%As expected, we find that the average family size of a $B$ allele decreases with distance into the $A$ side of the hybrid zone. 
% $F_B(m,x)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
%We find that at the selected locus, $F_B(m,x)$ decreases with x, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}). As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.

%One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection, leaving behind small families. In conjunction with block length, measuring $F$ can provide an additional signal for characterizing lineages strongly influenced by selection, by looking at levels of haplotype diversity in populations along a hybrid zone.

@@

\yaniv{remove this? or, what figure?}
Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short. This is measured by the mean length of the blocks adjacent to the block containing the selected site. A simple and intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond. Compared to other chromosomes in their new geographic location, these migrants will have on average longer $B$ haplotypes at the selected locus, and shorter $A$ haplotypes nearby. 
This finding suggests that, despite the fact that  haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.

@@

Our assumption of a single selected locus may have greater consequences. While our model is likely to be relevant to cases in which there are only a small number of targets of selection scattered throughout the genome, our predictions may differ significantly from situations in which the density of selected sites is higher. 
Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination. Both factors are expected to result in regions that are surrounded by even greater chunks of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection. 
 $s$ and relatively short map distances between linked incompatibilities)
 \yaniv{provide citation.}

@@

\yaniv{move this earlier}
Lastly, while our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms. This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.

>>>>

%%%%%%
\subsection*{The size of migrant families}

Within-ancestry haplotype diversity, i.e., the number of ancestral haplotypes of each type,
could provide additional information about whether introgression is through relatively few, successful migrants, 
or through many migrants that each contribute relatively little. 
In our simulations, the average local family size of a selected $B$ allele ($F_B(m,x)$) decreases with distance into the $A$ side of the hybrid zone, 
and is lowest far away from the zone center, where ancestry $B$ is at low frequency (Figure~\ref{Fig:family_size}). 
Unlinked loci have similar local family sizes similar to neutral simulations,
and loci linked to the selected locus have intermediate sizes.
This pattern is consistent with the prediction in our models that unfit lineages tend to be recent migrants, 
which will have smaller families.

%\alisa{add motivation-- additional information in haplotype diversity. Present alternate hypotheses. This ties in with our world view within our model}
%As expected, we find that the average family size of a $B$ allele decreases with distance into the $A$ side of the hybrid zone. 
% $F_B(m,x)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
%We find that at the selected locus, $F_B(m,x)$ decreases with x, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}). As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.
%One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection, leaving behind small families. In conjunction with block length, measuring $F$ can provide an additional signal for characterizing lineages strongly influenced by selection, by looking at levels of haplotype diversity in populations along a hybrid zone.

@@

\yb{remove this? or, what figure?}
Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short. This is measured by the mean length of the blocks adjacent to the block containing the selected site. A simple and intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond. Compared to other chromosomes in their new geographic location, these migrants will have on average longer $B$ haplotypes at the selected locus, and shorter $A$ haplotypes nearby. 
This finding suggests that, despite the fact that  haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.

@@

Our assumption of a single selected locus may have greater consequences. While our model is likely to be relevant to cases in which there are only a small number of targets of selection scattered throughout the genome, our predictions may differ significantly from situations in which the density of selected sites is higher. 
Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination. Both factors are expected to result in regions that are surrounded by even greater chunks of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection. 
 $s$ and relatively short map distances between linked incompatibilities)
 \yb{provide citation.}

@@

\yb{move this earlier}
Lastly, while our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms. This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.
